Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-07-16 review-responses.tex

  \begin{point}{}
    The inference of the deleterious DFE with GRAPES is quite inaccurate. 
    Is it possible that there is a factor of 2 error involved?
  \end{point}

  \reply{
As noted in the paper, conversion of selection inferences between tools is challenging.
We thus consulted with the developers of \polydfe and \grapes to confirm our conversions.
(The developer of \dadicli is an author on this paper.)
The inaccuracy of \grapes is a surprise to us.
Note that \grapes is also less accurate than \polydfe and \dadicli in inferring the shape of the distribution (Fig.~\ref{fig:homsap-dfe.ooa}B), which suggests the inaccuracy is not simply a scaling issue.
  }

  \begin{point}{\revref} % 21:
    Figure 3c does not have any error bars, it’s difficult to conclude if the inference is actually significantly different from the truth. 
    Which “N” was used to obtain estimates of “s” in Figure 3a? 
    This should be mentioned in the legend.  \end{point}

  \reply{
  %IG: I don't think that error bars are appropriate here and I tried to explain why
  The main purpose of the plots in Figure 3C is to show the actual differences between simulated
  and inferred distributions of $s$, since differences in mean and shape (shown in panels A-B) are not
  always very informative or insightful.
  To do this, we considered the average inferred DFE across the three replicate simualtions.
  We do not see much use in adding error bars to these plots reflecting the variation across the three replicates.
  Because of the somewhat noisy nature of this comparison, we only draw conclusions regarding qualitative differences
  between inferred and simulated distributions, namely,
  that  \polydfe and \dadicli tend to over-estimate the fraction of mutations under strong selection
  amd \grapes tends to over-estimate the fraction of mutations under weak selection.
  %IG: please check that this is correct
  Regarding the estimates of $\text{E}[s]$ shown in Figure 3A (and also Figure 4A and associated supplementary figures),
  these are typically values produced directly by the inference methods.
  In the last paragraph of the new methods section on DFE inference, we provide a complete description of the scaling
  we apply to the output of each method so that they are directly comparable to $s$ as simulated in \slim.
  }

### 2025-07-16 analysis2_manuscript.tex
some light edits/comments from dan

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulation replicates,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
        white bars represent the distribution used in the simulation. Standard error of the mean for each bin is shown in black. 
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

@@

    A second important finding is that inference of the DFE
    is sensitive to the dominance of deleterious mutations.
    All three of the DFE-inference methods applied here assume that deleterious mutations act
    in an additive fashion ($h=0.5$) by default.
    However, the simulated DFE for the vaquita genome included highly recessive mutations.
    Specifically, these simulations used a DFE where mutations with $s<-0.01$ are quite recessive,
    with $h<=0.01$, and where mutations with $-0.01<s<-0.001$ have $h=0.1$.
    We found that the three DFE-inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}A-B).
    This performance is not surprising since decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
    or an apparent decrease in the amount of selection.
    Thus, fitting a mis-specified DFE that assumes $h=0.5$ leads to underestimates of $E[s]$. \revpoint{1}{7}
    %IG: added this to respond to comment 1.7.
    %IG: I quote this in the response, so if we edit this, we should copy modified version to the response.
    Our analysis suggests that the distributions of $s$ estimated by these methods
    provide a reasonable approximation for the true distribution of $hs$ (Figure \ref{fig:vaquita-dfe}C).
    This points to a notable limitation of most existing DFE inference methods,
    which assume that selection coefficients are gamma-distributed.
    %DRS: Is the limitatio here that selection coefficients are gamma-distributed, or rather that h is fixed?
    As shown in Figure \ref{fig:vaquita-dfe}C,
    while a gamma distribution might fit the conditional distribution of $s$ given $h$, a relationship between $h$ and $s$ might result in a magrinal distribution for $hs$ which is not well approximated by a gamma distribution.

@@

%sweep power vs exon density
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8 \textwidth]{figures/HomSap/sims_runtime_plot.pdf}
    \caption{
    Runtime of simulated human chromosomes with and without background selection and demography.
    Each point represents a single simulation of a single chromosome under conditions labeled by facets.
    }
    \label{fig:sim-runtime}
\end{figure}

>>>>

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulation replicates,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated replicates (averaging the inferred gamma parameters);
        white bars represent the distribution used in the simulation. 
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

@@

    A second important finding is that inference of the DFE
    is sensitive to the dominance of deleterious mutations.
    All three of the DFE-inference methods applied here assume that deleterious mutations act
    in an additive fashion ($h=0.5$) by default.
    However, the simulated DFE for the vaquita genome included highly recessive mutations.
    Specifically, these simulations used a DFE where mutations with $s<-0.01$ are quite recessive,
    with $h<=0.01$, and where mutations with $-0.01<s<-0.001$ have $h=0.1$.
    We found that the three DFE-inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}A-B).
    This performance is not surprising since decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
    or an apparent decrease in the amount of selection.
    Thus, fitting a mis-specified DFE that assumes $h=0.5$ leads to underestimates of $E[s]$. \revpoint{1}{7}
    %IG: added this to respond to comment 1.7.
    %IG: I quote this in the response, so if we edit this, we should copy modified version to the response.
    Our analysis suggests that the distributions of $s$ estimated by these methods
    provide a reasonable approximation for the true distribution of $hs$ (Figure \ref{fig:vaquita-dfe}C).
    This points to a notable limitation of most existing DFE inference methods,
    which assume that selection coefficients are gamma-distributed.
    %DRS: Is the limitatio here that selection coefficients are gamma-distributed, or rather that h is fixed?
    As shown in Figure \ref{fig:vaquita-dfe}C,
    while a gamma distribution might fit the distribution of $s$, a relationship between $h$ and $s$ might result in
    $hs$ which is not well approximated by a gamma distribution.

@@
